The first high-redshift changing-look quasars

### 2020-07-10 tex/MNRAS_submission/rv1/CIV_CLQs_rv1.tex
another key BH_VirialMasses.py and CIV_CLQs_rv1 update

Both the quasar continuum (evaluated at 1450\AA), and the \civ\ line
luminosities follow a similar evolution, with a ratio of $\sim$20-30,
confirming that the main driver for emission line variability is
likely the broad band continuum itself.  For all sources except
J1638+2827, the slope of the continuum changes with luminosity
following a ``bluer-when-brighter'' pattern, suggesting that a
distinct emerging component is responsible for both the slope and
luminosity variations.  In J1638+2827 the opposite behaviour is
observed, especially in the first observation epoch.  However, this
may be a bias due to the limited wavelength range available which
extends to rest-frame $\lambda \sim 1240$~\AA\ for the first epoch,
while it extend to shorter wavelengths for the other epochs
(respectively ~\AA\ and 1010~\AA).  This suggests that the emerging
component is more prominent at UV wavelengths, and a sufficient
wavelength coverage is required to detect it. In all cases where the
\civ\ line profile is reliably constrained, we find that the \civ\
FWHM is approximately constant with maximum variations $\lesssim 1000$
km s$^{-1}$, despite significantly larger line luminosity variations.

@@

\subsection{Virial Black Hole Masses}\label{sec:BH_masses} 
The FWHMs of quasar broad lines are related to the mass of the SMBH
powering the quasar phenomenon, and that mass is assumed to be
constant on any human timescale.  Hence it is instructive to check
whether the virial products, which are the basic quantity used to
calculate single-epoch black hole mass estimates, show any
variation.

The continuum
luminosity provides a proxy for the BLR radius, and the broad-line
width (FWHM) provides a proxy for the virial velocity.  This ``virial
mass'' estimate is then expressed as:
{\bf
\begin{equation}
%  \log(M) = \log \left[c\left(  \frac{\nu L_\nu}{10^{42} {\rm erg\ s^{-1}}} \right)^{\gamma}  \left( \frac{\rm FWHM}{10^3 {\rm km\ s^{-1}}} \right)^2 \right] + {\rm constant}.
  \log \left( \frac{M_{\rm BH}}{M_{\odot}} \right)  = a + \log  \left[ \; \left(  \frac{L_{1450}} {10^{44} \; {\rm erg \; s^{-1}}}\right)^{\gamma} \, \right] + 2 \log \left( \frac{\rm {FWHM}}{{\rm km\; s^{-1}}} \right)
  \label{eqn:virial_masses}
\end{equation}
where $L_{1450}$ is the monochromatic (continuum) luminosity at 1450\AA\,
and FWHM is the full width at half-maximum of the broad \civ\ emission line.
The coefficient $\gamma$ and offset $a$ are generally deteremined empirically.

This approach assumes that the broad-line region (BLR) is virialized
\citep[e.g.][]{Shen2008b, Shen2011, Calderone2017,
Mejia-Restrepo2018}.  {\bf However, the \civ\ line is observed to have
line asymmetry due to outflows \citep[e.g.,][]{Gaskell1982}, very
strong blueshifts \citep[as mentioned in the Introduction, see
][]{Richards2011}, and a component of the line profile that does not
reverberate \citep[][]{Denney2012}.  These issues make the virial BH
mass estimate uncertain. \citet{Runnoe2013}, 
\citet{Mejia-Restrepo2016}, \citet{Coatman2017}, 
\citet{Mejia-Restrepo2018} and \citet{Grier2019} all present studies
into improving \civ-based single-epoch black hole mass estimates. With
these cavaets in mind, we acknowledge the BH mass estimates will have
a large error, likely $\sim$0.5 dex. }

We report a virial product for the \civ\ CLQs in the last column of
Table~\ref{tab:QSFit-results}.  {\bf Since our purpose is to check if
this measurement varies significantly across epochs, here we use
$\gamma=0.5$ as the power-law slope and are not directly concerned
with the offset.  The quoted uncertainties account only for the
propagation of uncertainties in the continuum luminosities and line
widths.  As noted above, with typical uncertainties associated to
single epoch virial mass around $\sim$0.5 dex,} this implies that the
virial products for our three high-redshift CLQs are all remarkably
constant and compatible with single black hole masses per source, even
in those cases where the \civ\ estimates were deemed potentially
unreliable. The object showing the largest variation is J2228+2201,
with the BH mass estimate spanning a range of 0.66~dex.

\citet{VestergaardPeterson2006} have $\gamma=0.53$ and the offset as $a=0.660$, 
which is widely adopted \citep[e.g.,][]{Chen2019, Yao2019}.
\citet{Kozlowski2017} have similar values \citet{VestergaardPeterson2006} , with $(a,\gamma)=(0.64, 0.53)$,
but note there is a problem with the FWHMs for \civ\ for those quasars that are reported in
both the DR7 Quasar catalog and DR12 Quasar catalog.
\citet{Grier2019} measure a slope of $\gamma=0.51\pm0.05$ for their sample of high-redshift
$z>1.3$ quasars that show significant lag measurements,
and  $\gamma=0.52\pm0.04$ for their ``gold sample'' of 16 of
their highest-confidence objects. }

@@

From the virial mass estimates, all our objects have SMBH masses $M_{\rm BH} >10^{9} M_{\odot}$.  This is at the upper end of SMBH masses at all epochs, and towards the extreme of the mass distribution for $z\sim2$ quasars.  Figure~\ref{fig:CIV_MBHvsREW} compares the luminosity \civ\ EW versus the virial SMBH masses for our multi-epoch 
observations to a sample of $\approx$20,000 $z>1.5$ SDSS quasars from
the QSFit catalog \citep{Calderone2017}.

{\bf With the black hole masses in hand, we can calculate the Eddington luminosities of 
the three quasars. To be consistent with 
For the Bolometric luminosity, $L_{\rm Bol}$, \citet{Shen2011} and 
\citet{Kozlowski2017} use the bolometric corrections of BC$_{1350}$ = 3.81 
from the composite spectral energy distribution (SED) in \citet{Richards2006b}.

We report the black hole \civ\ virial masses, the bolometric luminosities
and the Eddington ratios ($\eta=\log(L_{}/L_{Edd})$ in Table~\ref{tab:Eddington_ratios}. 
}

%We use the mean of five values --- three from our three epochs and two from \citet{Shen2011} --- to calculate an adopted SMBH mass of $\log (M_{\rm BH, vir}/M_{\odot}) = 9.73$ for J1205+3422.  J1638+2827 also has two virial mass measurements from \citet{Shen2011}, as well as two measurements from \citet{Kozlowski2017}. The mean mass measurement adopted for J1638+2827 is $\log(M_{\rm BH, vir}/M_{\odot} = 9.09$. For J2228+2201, there are two \citet{Kozlowski2017} measurements, leading to a mean adopted mass of $\log(M_{\rm BH, vir}/M_{\odot} = 9.37$. We use these mean SMBH masses when calculating the Eddington ratios.

@@

%%
%%  These values are generated using the 
%%
\begin{table}
  \centering
  \begin{tabular}{l l  cc         r}
    \hline
    \hline
                      &                  &                                      &                              & \\
Object            & MJD           &  $\log (M/M_{\odot})$  & log($L_{\rm bol}$)  &  $\eta_{\rm Edd}$  \\
                      &                  &                                      &                              & \\
    \hline
                      & $^{\dagger }$53498  &  9.49              &  47.22                  &  -0.378   \\
                      & 53498                      &    9.39            &                    &  $-$0.61   \\
 J1205+3422 & 58538                       &    9.78           &                      &   $-$1.76 \\  
                     & 58693                         &    9.38          &                   &  $-$1.26  \\                    
    \hline
                     & $^{\dagger }$54553    &  8.74            &  46.17   & $-$0.677  \\
                     & 54553                         &   8.84                 &             & $-$1.03  \\
J1638+2827  & $^{\ddagger }$55832  &    9.13                  & 45.98   & $-$1.21  \\      
                      & 55832                        &   8.80                  & 45.98   & $-$1.21  \\     
                      & 58583                        &   9.01                  & 46.50   & $-$0.69   \\
    \hline
                         & $^{\ddagger }$56189   & 8.73                  & 46.23  &  $-$1.24 \\
                         & 56189                         & 8.70                  & 46.23  &  $-$1.24 \\
J2228+2201     & 56960                          &  9.39               & 47.34  & $-$0.12  \\      
                         & 58693                         &  8.97               & 46.83   &  $-$0.64 \\
    \hline
    \hline
  \end{tabular}
  \caption{The \civ\ virial mass estimates, bolometric luminosities and Eddington ratios for the
    three quasars.
    The virial masses are calculated using the measured monochromatic luminosites and FWHM
    $\dagger$From the \citet{Shen2011} SDSS DR7 Quasar catalog.
    $\ddagger$From the \citet{Kozlowski2017} SDSS-III BOSS DR12 quasar catalog.
%    are given in either \citet{Shen2011} or    \citet{Kozlowski2017} for the three CLQs.
    We scale these values using
    our measured $\nu L_{\nu}$ continuum luminosity values.  From
    Section~\ref{sec:BH_masses}, we assume: $\log(M_{\rm BH,
      vir}/M_{\odot}) = 9.726$ for J1205+3422; $\log(M_{\rm BH,
      vir}/M_{\odot}) = 9.091$ for J1638+2827, and $\log(M_{\rm BH,
      vir}/M_{\odot}) = 9.366$ for J2228+2201.  The Eddington luminosity
    $L_{\rm Edd} = 1.26\times10^{38} (M_{\rm BH}/M_{\odot})$ erg s$^{-1}$
    and $\eta_{\rm Edd}$ is the log of the Eddington ratio.}
\label{tab:Eddington_ratios} 
\end{table}

@@

\subsection{Continuum and Line Changes: Comparisons to Theoretical Expectations}\label{sec:theory}
The \civ\ line is one of the strongest collisionally excited lines in
quasar spectra \citep[e.g.,][]{HamannFerland1999}.
{\bf Early measurements from e.g., 
  \citet{PoggePeterson1992} and \citet{Peterson1997book}, 
  using the well-studied Seyfert galaxy NGC 5548,
  suggested the location of \civ\ emission was 
%  probes the photoionisation environment
  produced by the innermost disk, with the line emission
  responding to the underlying continuum emission in only
  a matter of $\sim$ 8 days. 
as indicated by reverberation mapping time-delay measurements.
%{\bf (a) I would recommend citing some representative papers to support the statement made in the first sentence of section 4.2 about the lag of the C IV line. The most appropriate papers are those dealing with quasars of comparable luminosity (and probably comparable redshift) as the quasars studied here. I am somewhat skeptical of this statement because the papers that I consulted.}
However, more recent measurements from e.g. \citet[][and references therein]{Grier2019},
suggest that the lag between the \civ\ line and the continuum are of order $>100$ observed-frame days. 
which would place the \civ emitting region far from the inner accretion disk.
Importantly, many quasars in the \citet{Grier2019} sample are of
comparable redshift and luminosity to SDSS J1205+3422, J1638+2827, and J2228+2201.}

@@

\citet{NodaDone2018} note that the soft X-ray excess -- the excess of
X-rays below 2 keV with respect to the extrapolation of the hard X-ray
spectral continuum model, and a common feature among Type~1 AGN --
produces most of the ionizing photons, so its dramatic drop can lead
to the disappearance of the broad-line region, driving the
changing-look phenomena. \citet{NodaDone2018} go on to make a
connection between state changes in Galactic binaries and CLQs.
%
In Galactic binaries spectral hardening
corresponds to the soft-to-hard state transition in black hole
binaries, which occurs at $L / L_{\rm Edd} \sim 0.02$ (i.e.,
$\eta_{\rm Edd} \sim -1.7$).  During this state transition, the inner
disc evaporates into an advection-dominated accretion flow, while the
overall drop in luminosity appears consistent with the hydrogen
ionisation disc instability.  \citet{NodaDone2018} predict that
changing-look AGNs are similarly associated with state transitions at
$L / L_{\rm Edd}$ of a few percent.

>>>>

Both the quasar continuum (evaluated at 1450\AA), and the \civ\ line
luminosities follow a similar evolution, with a ratio of $\sim$20-30,
confirming that the main driver for emission line variability is
likely the broad band continuum itself.  For all sources except
J1638+2827, the slope of the continuum changes with luminosity
following a ``bluer-when-brighter'' pattern, suggesting that a
distinct emerging component is responsible for both the slope and
luminosity variations.  In J1638+2827 the opposite behaviour is
observed, especially in the first observation epoch.  However, this
may be a bias due to the limited wavelength range available which
extends to rest-frame $\lambda$$\sim$1240~\AA\ for the first epoch,
while it extend to shorter wavelengths for the other epochs
(respectively 1100~\AA\ and 1010~\AA).  This suggests that the emerging
component is more prominent at UV wavelengths, and a sufficient
wavelength coverage is required to detect it. In all cases where the
\civ\ line profile is reliably constrained, we find that the \civ\
FWHM is approximately constant with maximum variations $\lesssim$1000 
km s$^{-1}$, despite significantly larger line luminosity variations.

@@

\subsection{Virial Black Hole Masses}\label{sec:BH_masses} 
The FWHMs of quasar broad lines are related to the mass of the SMBH
powering the quasar phenomenon, and that mass is assumed to be
constant on any human timescale.  Hence it is instructive to check
whether the virial products, which are the basic quantity used to
calculate single-epoch black hole mass estimates, show any
variation. The continuum luminosity provides a proxy for the BLR
radius, and the broad-line width (FWHM) provides a proxy for the
virial velocity. This approach assumes that the broad-line region
(BLR) is virialized \citep[e.g.][]{Shen2008b, Shen2011, Calderone2017,
Mejia-Restrepo2018} and the virial mass estimate is then:
{\bf
\begin{equation}
%  \log(M) = \log \left[c\left(  \frac{\nu L_\nu}{10^{42} {\rm erg\ s^{-1}}} \right)^{\gamma}  \left( \frac{\rm FWHM}{10^3 {\rm km\ s^{-1}}} \right)^2 \right] + {\rm constant}.
  \log \left( \frac{M_{\rm BH}}{M_{\odot}} \right)  = a + \log  \left[ \; \left(  \frac{L_{1450}} {10^{44} \; {\rm erg \; s^{-1}}}\right)^{\gamma} \, \right] + 2 \log \left( \frac{\rm {FWHM}}{{\rm km\; s^{-1}}} \right)
  \label{eqn:virial_masses}
\end{equation}
where $L_{1450}$ is the monochromatic continuum luminosity at 1450\AA\,
and FWHM is the full width at half-maximum of the broad \civ\ emission line.
The coefficient $\gamma$ and offset $a$ are generally deteremined empirically.

 {\bf The \civ\ line is observed to have
line asymmetry due to outflows \citep[e.g.,][]{Gaskell1982}, very
strong blueshifts \citep[as mentioned in the Introduction, see
][]{Richards2011}, and a component of the line profile that does not
reverberate \citep[][]{Denney2012}.  These issues make the virial BH
mass estimate uncertain. \citet{Runnoe2013}, 
\citet{Mejia-Restrepo2016}, \citet{Coatman2017}, 
\citet{Mejia-Restrepo2018} and \citet{Grier2019} all present studies
into improving \civ-based single-epoch black hole mass estimates. With
these cavaets in mind, we acknowledge the BH mass estimates will have
a large error, likely $\sim$0.5 dex. }

We report the virial product for the \civ\ CLQs in the last column of
Table~\ref{tab:QSFit-results}.  {\bf Since our purpose is to check if
this measurement varies significantly across epochs, here we use
$\gamma=0.5$ as the power-law slope and are not directly concerned
with the offset.  The quoted uncertainties account only for the
propagation of uncertainties in the continuum luminosities and line
widths.  As noted above, with typical uncertainties associated to
single epoch virial mass around $\sim$0.5 dex,} this implies that the
virial products for our three high-redshift CLQs are all remarkably
constant and compatible with single black hole masses per source, even
in those cases where the \civ\ estimates were deemed potentially
unreliable. The object showing the largest variation is J2228+2201,
with the BH mass estimate spanning a range of 0.66~dex.

For Equation~\ref{eqn:virial_masses}., \citet{VestergaardPeterson2006}
measure $\gamma=0.53$ and the offset as $a=0.660$, and these estimates
are widely adopted \citep[e.g.,][]{Shen2011, Chen2019, Yao2019}.
\citet{Kozlowski2017} have similar values \citet{VestergaardPeterson2006} , with $(a,\gamma)=(0.64, 0.53)$,
but note there is a problem with the FWHMs for \civ\ for those quasars that are reported in
both the DR7 Quasar catalog and DR12 Quasar catalog.
\citet{Grier2019} measure a slope of $\gamma=0.51\pm0.05$ for their sample of high-redshift
$z>1.3$ quasars that show significant lag measurements,
and  $\gamma=0.52\pm0.04$ for their ``gold sample'' of 16 of
their highest-confidence objects. }

@@

From the virial mass estimates, all our objects have SMBH masses $M_{\rm BH} \approx 500\times10^{8} M_{\odot}$.  This is at the upper end of SMBH masses at all epochs, and towards the extreme of the mass distribution for $z\sim2$ quasars.  Figure~\ref{fig:CIV_MBHvsREW} compares the luminosity \civ\ EW versus the virial SMBH masses for our multi-epoch 
observations to a sample of $\approx$20,000 $z>1.5$ SDSS quasars from
the QSFit catalog \citep{Calderone2017}.

{\bf With the black hole masses in hand, we can calculate the
Eddington luminosities of the three quasars. To be consistent with For
the Bolometric luminosity, $L_{\rm Bol}$, \citet{Shen2011} and
\citet{Kozlowski2017} use the bolometric corrections of BC$_{1350}$ =
3.81 (from the composite spectral energy distribution (SED) in
\citet{Richards2006b}) and to be consistent with those measurements we
adopt the same correction.  We report the the bolometric luminosities
and the Eddington ratios, $\eta=\log(L_{}/L_{Edd})$, in
Table~\ref{tab:Eddington_ratios}.  }

@@

%%
%%  These values are generated using the 
%%
\begin{table}
  \centering
  \begin{tabular}{l c cc l}
    \hline
    \hline
                      &                 &                                     &                               & \\
Object            & MJD          &  $\log (M/M_{\odot})$  & log($L_{\rm bol}$)    &  $\eta_{\rm Edd}$  \\
                      &                 &                                     &                               & \\
    \hline
\multirow{4}{*}{J1205+3422} & 53498          &   9.49$^{\dagger }$   &   47.22$^{\dagger }$  & $-$0.38$^{\dagger }$   \\
                                             &  53498        &   9.39                       &   47.20                      & $-$0.61   \\
                                             & $^{*}$58538  &   9.78                       &  46.03                      & $-$1.84 \\  
                                             & 58693            &  9.38                       &  46.54                      &  $-$0.95  \\                    
    \hline
\multirow{5}{*}{J1638+2827} & 54553            & 8.74$^{\dagger }$     &  46.17$^{\dagger }$   & $-$0.68$^{\dagger }$  \\
                                              & 54553            & 8.84                        &   46.19    & $-$0.75  \\
                                             & 55832           & 9.13$^{\ddagger }$  &    46.72    & $-$0.72$^{\ddagger }$  \\      
                                             & 55832          & 8.80                         &   45.99    & $-$0.91 \\     
                                             & 58583           & 9.01                        &    46.52    & $-$0.60   \\
    \hline
\multirow{4}{*}{ J2228+2201} & 56189         & 8.73$^{\ddagger }$ & 46.23$^{\ddagger }$  &  $-$1.32$^{\ddagger }$ \\
                                              &  56189      & 8.70                       & 45.51                        &  $-$1.29 \\
                                              & 56960         &  9.39                      &  46.54                        & $-$0.95  \\      
                                              & 58693          &  8.97                      & 46.03                         &  $-$1.05 \\
    \hline
    \hline
  \end{tabular}
  \caption{The \civ\ virial mass estimates, bolometric luminosities and Eddington ratios for the
    three quasars.
    The virial masses are calculated using the measured 1450\AA\ monochromatic luminosites and \civ\ FWHM
    as reported by QSFit, with $^{*}$ designating low SNR line measurements.
        $\dagger$Values from the \citet{Shen2011} SDSS DR7 Quasar catalog.
    $\ddagger$Values from the \citet{Kozlowski2017} SDSS-III BOSS DR12 quasar catalog.
    The Eddington luminosity is 
    $L_{\rm Edd} = 1.26\times10^{38} (M_{\rm BH}/M_{\odot})$ erg s$^{-1}$
    and $\eta_{\rm Edd}$ is the log of the Eddington ratio.}
\label{tab:Eddington_ratios} 
\end{table}

@@

\subsection{Continuum and Line Changes: Comparisons to Theoretical Expectations}\label{sec:theory}
The \civ\ line is one of the strongest collisionally excited lines in
quasar spectra \citep[e.g.,][]{HamannFerland1999}.  {\bf Early
measurements from e.g., \citet{PoggePeterson1992} and
\citet{Peterson1997book}, using the well-studied Seyfert galaxy NGC
5548, suggested the location of \civ\ emission was produced by the
innermost disk, with the line emission responding to the underlying
continuum emission in only a matter of $\sim$8 days.  as indicated by
reverberation mapping time-delay measurements.  However, more recent
measurements from e.g. \citet[][and references therein]{Grier2019},
suggest that the lag between the \civ\ line and the continuum are of
order $>100$ observed-frame days.  which would place the \civ emitting
region far from the inner accretion disk.  Importantly, many quasars
in the \citet{Grier2019} sample are of comparable redshift and
luminosity to SDSS J1205+3422, J1638+2827, and J2228+2201.}

@@

\citet{NodaDone2018} note that the soft X-ray excess -- the excess of
X-rays below 2 keV with respect to the extrapolation of the hard X-ray
spectral continuum model, and a common feature among Type~1 AGN --
produces most of the ionizing photons, so its dramatic drop can lead
to the disappearance of the broad-line region, driving the
changing-look phenomena. \citet{NodaDone2018} go on to make a
connection between state changes in Galactic binaries and CLQs.  In
Galactic binaries spectral hardening corresponds to the soft-to-hard
state transition in black hole binaries, which occurs at $L / L_{\rm
Edd} \sim 0.02$ (i.e., $\eta_{\rm Edd} \sim -1.7$).  During this state
transition, the inner disc evaporates into an advection-dominated
accretion flow, while the overall drop in luminosity appears
consistent with the hydrogen ionisation disc instability.
\citet{NodaDone2018} predict that changing-look AGNs are similarly
associated with state transitions at $L / L_{\rm Edd}$ of a few
percent.
